Algebraic Principles for Concurrency Verification Tools

### 2013-10-08 Paper/paper.tex
Figured out rules for RG algebra w/ operator

Axiom (\ref{rg1}) states that interference from the a rely condition
being run twice in parallel is no different from just the
interference from that rely condition. Axiom (\ref{rg2}) states that
interference from a single rely condition is less than interference
from itself and another rely condition. Axiom (\ref{rg3}) allows a
rely or guarantee to be split across sequential programs. Axiom
(\ref{rg4}) is similar to Axiom (\ref{rg3}) in intent, except it deals
with finite iteration.

@@

Jones quintuples can be encodeded in this setting as
\begin{align}
r, g \vdash \{p\} c \{q\} \iff p(r\|c) \le q \land c \le g.
\end{align}
With this encoding we can derive the standard rely-guarantee inference rules, as
shown in Figure \ref{fig:rgrules}.

@@

By using residuals quintuples can be encoded in the following
equivalent way,
\begin{align}
r, g \vdash \{p\} c \{q\} \iff c \le r/(p \rightarrow q) \sqcap g \label{refine}.
\end{align}
This encoding allows us to think in terms of program
refinement, as $r/(p \rightarrow q) \sqcap g$ defines the weakest
program that when placed in parallel with interference from $r$, and
guaranteeing interference at most $g$, goes from $p$ to $q$.

@@

While this setting is adequate for deriving the standard inference
rules, the equality is too strong to capture many interesting
statements about concurrent programs. Consider the statement,
$x = y \implies x\|z = y\|z$,
which holds in our algebra and can be read as follows; if $x$ and $y$ are
equal, then they must be equal under all possible interference from
any arbitrary $z$. This would seem to preclude any reasoning about
interference in a fine-grained way, but we can work around this by
using an \emph{interior operator}. In a rely-guarantee algebra an
interior operator is any function $f : K \to K$ which satisfies
\begin{align*}
f(f(x)) &= f(x) \tag{$f$ is idempotent}\\
x \le y \implies f(x) &\le f(y) \tag{$f$ is isotone}\\
f(x) &\le x \tag{$f$ is coextensive}
\end{align*}

>>>>

Axiom (\ref{rg1}) states that interference from a rely condition being
run twice in parallel is no different from just the interference from
that rely condition begin run once in parallel. Axiom (\ref{rg2})
states that interference from a single rely condition is less than
interference from itself and another rely condition. Axiom (\ref{rg3})
allows a rely or guarantee to be split across sequential
programs. Axiom (\ref{rg4}) is similar to Axiom (\ref{rg3}) in intent,
except it deals with finite iteration.

@@

Jones quintuples can be encodeded in this setting as
\begin{align}
r, g \vdash \{p\} c \{q\} \iff p(r\|c) \le q \land c \le g. \label{quin}
\end{align}
With this encoding we can derive the standard rely-guarantee inference
rules, as shown in Figure \ref{fig:rgrules}. Thus equations
(\ref{rg1}) to (\ref{rg4}), which are all necessary to derive these
rules, somehow represent a minimal set of axioms from which these
inference rules can be derived.

@@

By using residuals quintuples can be encoded in the following way,
which is equivalent to the encoding in Equation (\ref{quin}).
\begin{align}
r, g \vdash \{p\} c \{q\} \iff c \le r/(p \rightarrow q) \sqcap g \label{refine}.
\end{align}
This encoding allows us to think in terms of program refinement, as
$r/(p \rightarrow q) \sqcap g$ defines the weakest program that when
placed in parallel with interference from $r$, and guaranteeing
interference at most $g$, goes from $p$ to $q$---which is essentially
a generic specification for a concurrent program.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Breaking Compositionality}
\label{sec:INT}

While the algebra in section \ref{sec:RG} is adequate for deriving the
standard inference rules, the equality is too strong to capture many
interesting statements about concurrent programs. Consider the
congruence rule for parallel composition, which is inherent in the
algebraic approach:
\begin{align*}
x = y \implies x\|z = y\|z.
\end{align*}
This can be read as follows; if $x$ and $y$ are equal, then they must
be equal under all possible interference from any arbitrary $z$. At
first glance this might seem to preclude any reasoning about
interference in a fine-grained way using only algebra. This is not the
case, but breaking such inherent compositionality in just the right way to
capture interesting properties of interference requires some extra work.

The most obvious way we can acheive this is by simply expanding our
rely-guarantee algebra with an arbitrary function $f$. And redefining our quintuples
as,
\begin{align*}
r, g \vdash \{p\} c \{q\} \iff f(p(r\|c)) \le f(q) \land c \le g.
\end{align*}

 but we can work around this by
using an \emph{interior operator}. In a rely-guarantee algebra an
interior operator is any function $f : K \to K$ which satisfies
\begin{align*}
f(f(x)) &= f(x) \tag{$f$ is idempotent}\\
x \le y \implies f(x) &\le f(y) \tag{$f$ is isotone}\\
f(x) &\le x \tag{$f$ is coextensive}
\end{align*}
